\textbf{Teaching a machine to generate an artificial universe}

### 2020-08-03 report.tex
formatted report.tex for github viewing

\title{\textbf{Teaching a machine to generate an artificial universe}}
\author[1]{Jed Homer\thanks{}}
\author[1]{Christopher Messenger}
\author[1]{Martin Hendry}
\affil[1]{Institute for Gravitational Research, School of Physics and Astronomy, University of Glasgow, G12 8QQ, UK}
%Institute for Gravitational Research, 
\begin{document}
\date{}
%\lhead{\textbf{Version $2.7$}}
\twocolumn[
	\begin{@twocolumnfalse}
		\maketitle
		\begin{abstract}
		    This work presents separate two- and three-dimensional generative frameworks that use Generative Adversarial Networks to create realistic samples of cosmological structure. Provisional samples of the virtual cosmic web are obtained from N-body simulation data that is restricted to volumes and projections of size $80 \, h^{-1} \, \text{Mpc}$. The generated samples are shown to demonstrate a high degree of coherence with their simulated counterparts and the distinction cannot be made visually by human experts. The nature of the separation between generated and N-body simulation samples is quantified with repeated tests of the Kullback-Leibler divergence on ensembles of samples alongside samples with uniformly distributed pixel densities. It is found that the implementations secure a coherence to within the same order of magnitude as a self comparison of the simulation samples. The main advantage of the generative approach to simulating large volumes of cosmological structure is that much less time is required than creating N-body simulations on high performance computing resources. Future explorations of the method that are relevant to the next paradigm in cosmological simulations are also outlined. The most pertinent application being for a generator that can produce samples that maintain high statistical coherence to simulation data at various redshifts. At this stage, the generative artificial intelligence can then be said to be capable of generating an artificial universe.

            %\textit{Nbody data, GANs, coherence, main results, implications and amplications, }

@@

% \subsubsection{2D-GAN implementation} % DESCRIBE SPECIFICS OF THIS IMPLEMENTATION
% \textit{Architecture} --- 
% The networks follow the structuring of the DCGAN models~\cite{dcgan} and are written in Keras~\cite{keras}. A full 
% description of the layers in the 2D-GAN architecture is given in Table~\ref{table:2dgan_archi} of 
% Appendix~\ref{appendix:gan_setups}. The generator transforms a 100-dimensional vector drawn from a Gaussian prior 
% $\mathbf{z}\sim \mathcal{N} (0,1)$ into a $32\times 32$ single-channel image. The latent-space vector is followed by a 
% fully connected layer that is reshaped to a stack of feature maps for the two convolutional layers. Batch-normalisation 
% is only used after the fully connected layer. Through each layer hyperbolic tangent activations are used. See Appendices~\ref{appendix:app_acts} and~\ref{app_BN} for more detail on batch-normalisation and activation functions.

@@

% The differences between the training-uniform and generated-uniform means and variances are similar to the same values in 
% the 2D-comparisons. These values show the significance of the training-generated comparisons in the top row of Figure~\ref{fig:3dgan_kltrials}, which are a nearly a whole order of magnitude smaller. The mean $D_{KL}$ values of these 
% comparisons are almost at the value of the training-training comparison means which is more evidence for the understandings
% of the training data by the generator. The similarity shown by the $D_{KL}(t||g)$ and $D_{KL}(g||t)$ values may mostly be 
% due to the similar pixel density distributions of the training and generated data. Considering that the lack of filament 
% structure in the larger scales of the generated data compared to the training data is not shown in the $D_{KL}$ comparisons,
% this may be true. The pixel density distribution is an important characteristic of the images but the just obtaining the 
% similarity between the $t$ and $g$ distrubitions with this property would not guarantee that all the structure in the 
% training data was present. There are many combinations of pixels that make up a matching pixel-density distribution for 
% two histogram distributions, but far less of those show the structure of the cosmic web. The higher standard deviations 
% in the $D_{KL}$ histograms for the 2D case in all the kernel sizes compared to the 3D cases is due to the larger size of 
% features relative to total size of the image that shows them in either case.

>>>>

\title{\textbf{Teaching a machine to generate an artificial universe}}
\author[1]{Jed Homer\thanks{}}
\author[1]{Christopher Messenger}
\author[1]{Martin Hendry}
\affil[1]{Institute for Gravitational Research, School of Physics and Astronomy, University of Glasgow, G12 8QQ, UK}

\begin{document}
\date{}
%\lhead{\textbf{Version $2.7$}}

\twocolumn[
	\begin{@twocolumnfalse}
		\maketitle
		\begin{abstract}
		    This work presents separate two- and three-dimensional generative frameworks that use Generative Adversarial 
		    Networks to create realistic samples of cosmological structure. Provisional samples of the virtual cosmic web 
		    are obtained from N-body simulation data that is restricted to volumes and projections of size $80 \, h^{-1} 
		    \, \text{Mpc}$. The generated samples are shown to demonstrate a high degree of coherence with their simulated
		    counterparts and the distinction cannot be made visually by human experts. The nature of the separation between
		    generated and N-body simulation samples is quantified with repeated tests of the Kullback-Leibler divergence on
		    ensembles of samples alongside samples with uniformly distributed pixel densities. It is found that the 
		    implementations secure a coherence to within the same order of magnitude as a self comparison of the simulation
		    samples. The main advantage of the generative approach to simulating large volumes of cosmological structure is 
		    that much less time is required than creating N-body simulations on high performance computing resources. 
		    Future explorations of the method that are relevant to the next paradigm in cosmological simulations are also 
		    outlined. The most pertinent application being for a generator that can produce samples that maintain high 
		    statistical coherence to simulation data at various redshifts. At this stage, the generative artificial 
		    intelligence can then be said to be capable of generating an artificial universe.

            %\textit{}

@@

% \subsubsection{2D-GAN implementation} % DESCRIBE SPECIFICS OF THIS IMPLEMENTATION
% \textit{Architecture} --- 
% The networks follow the structuring of the DCGAN models~\cite{dcgan} and are written in Keras~\cite{keras}. A full 
% description of the layers in the 2D-GAN architecture is given in Table~\ref{table:2dgan_archi} of 
% Appendix~\ref{appendix:gan_setups}. The generator transforms a 100-dimensional vector drawn from a Gaussian prior 
% $\mathbf{z}\sim \mathcal{N} (0,1)$ into a $32\times 32$ single-channel image. The latent-space vector is followed by a 
% fully connected layer that is reshaped to a stack of feature maps for the two convolutional layers. Batch-normalisation 
% is only used after the fully connected layer. Through each layer hyperbolic tangent activations are used. See 
% Appendices~\ref{appendix:app_acts} and~\ref{app_BN} for more detail on batch-normalisation and activation functions.

@@

% The differences between the training-uniform and generated-uniform means and variances are similar to the same values in 
% the 2D-comparisons. These values show the significance of the training-generated comparisons in the top row of 
% Figure~\ref{fig:3dgan_kltrials}, which are a nearly a whole order of magnitude smaller. The mean $D_{KL}$ values of these 
% comparisons are almost at the value of the training-training comparison means which is more evidence for the understandings
% of the training data by the generator. The similarity shown by the $D_{KL}(t||g)$ and $D_{KL}(g||t)$ values may mostly be 
% due to the similar pixel density distributions of the training and generated data. Considering that the lack of filament 
% structure in the larger scales of the generated data compared to the training data is not shown in the $D_{KL}$ comparisons,
% this may be true. The pixel density distribution is an important characteristic of the images but the just obtaining the 
% similarity between the $t$ and $g$ distrubitions with this property would not guarantee that all the structure in the 
% training data was present. There are many combinations of pixels that make up a matching pixel-density distribution for 
% two histogram distributions, but far less of those show the structure of the cosmic web. The higher standard deviations 
% in the $D_{KL}$ histograms for the 2D case in all the kernel sizes compared to the 3D cases is due to the larger size of 
% features relative to total size of the image that shows them in either case.
